---
title: "Zoom browser"
author: "Myles Lewis"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 2
    number_sections: true
vignette: >
  %\VignetteIndexEntry{locuszoomr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
library(locuszoomr)
ensdb_eval <- requireNamespace("EnsDb.Hsapiens.v75", quietly = TRUE)
```

Here we demonstrate how to use the zoom browser to interactively browse whole
genome GWAS or eQTL data.

# Example GWAS data

Example whole genome summary statistics from an SLE GWAS study can be downloaded
from [EBI](https://www.ebi.ac.uk/gwas/studies/GCST003156) and loaded as follows.

```{r eval=FALSE}
# FTP download full summary statistics (308 MB) from SLE GWAS from
# https://www.ebi.ac.uk/gwas/studies/GCST003156
library(data.table)
SLE_gwas <- fread('../bentham_2015_26502338_sle_efo0002690_1_gwas.sumstats.tsv')
```

The zoom browser is started using `zoom()`. Like the `locus()` function,
`zoom()` tries to autodetect the columns for SNP id, chromosome, position and
p-value. If you have more complex column names, then the arguments `chrom`,
`pos`, `p` and `labs` can be used to specify the appropriate columns in your
dataset.

An appropriate hg19 or hg38 (or other species) ensembl database of
genome information must be specified. In the example below, the GWAS data is
hg19.

```{r eval=FALSE}
zoom(SLE_gwas, ens_db = "EnsDb.Hsapiens.v75")
```

# Add recombination rate

Recombination rate can be shown on a secondary *y* axis. Whole recombination
rate track data file (around 30 MB) should be downloaded from UCSC genome
browser (documented
[here](https://genome.ucsc.edu/cgi-bin/hgTrackUi?g=recombRate2)).

The download site can be accessed at
https://hgdownload.soe.ucsc.edu/gbdb/hg38/recombRate/. For hg38, download
`recomb1000GAvg.bw`.

For hg19, the link is https://hgdownload.soe.ucsc.edu/gbdb/hg19/decode/ and the
default track we use is `hapMapRelease24CombinedRecombMap.bw`.

The .bw track file can then be loaded into R as a `GRanges` object using
`import.bw()`, and then used directly by `zoom()`.

```{r eval=FALSE}
library(rtracklayer)
recomb.hg19 <- import.bw("/../hapMapRelease24CombinedRecombMap.bw")

zoom(SLE_gwas, ens_db = "EnsDb.Hsapiens.v75", recomb = recomb.hg19)
```

# Citation

If you use this package please cite as:

Lewis MJ, Wang S. (2025) locuszoomr: an R package for visualising
publication-ready regional gene locus plots. *Bioinformatics Advances* 2025;
vbaf006, [doi:10.1093/bioadv/vbaf006](https://doi.org/10.1093/bioadv/vbaf006)
